---
title: "preprocess_for_cross-species"
author: "Beth Moore"
date: "2023-07-07"
output: html_document
---
# Preprocessing for cross species analysis
# to have cross species, we need ortholog genes
# we then need to subset both reference and query by these genes
# reference data will also be subseted by its metadata

# Load libraries, set wd, and source functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(patchwork)
library(scran)
library(BiocParallel)
library(DropletUtils)
library(cowplot)
library(harmony)
library(SoupX)
library(scDblFinder)
library(reticulate)
library(purrr)
library(jsonlite)
library(rmarkdown)
use_condaenv("PIPELINE_1.03", conda = "/isiseqruns/jfreeman_tmp_home/bin/miniconda3/condabin/conda")
setwd("/w5home/bmoore/scRNAseq/GAMM/human_data/reh_cellrep_2020/")
timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
output <- paste0("output_", timestamp)
dir.create(output, showWarnings = FALSE)
file.copy("/w5home/bmoore/scRNAseq_library/config.json", file.path(output, "config.json"))
config <- fromJSON(file.path(output, "config.json"))
output <- paste0(output, "/")
source("/w5home/bmoore/scRNAseq_library/src/sc_pipeline_functions.R")
```
# Read in data
```{r data}
exp.matrix <- read_aligned_data("~/scRNAseq/GAMM/human_data/reh_cellrep_2020/output_reh/Solo.out/GeneFull/","human_ref")
# load gamm pig data sample 2
gamm.data <- read_aligned_data("/w5home/bmoore/scRNAseq/GAMM/output_S2_mm_mt/Solo.out/GeneFull/","gamm.data.S2")
```

# run soupx
```{r soupx}
all_aligned_data <- list(exp.matrix,gamm.data)
soupX_objs <- purrr::map(all_aligned_data, ~prep_seurat_and_soupX(data.raw = .x$raw, data = .x$filtered, project = .x$project), .names = purrr::map_chr(all_data, 'project'))
# write out data
DropletUtils:::write10xCounts("reh_soupX_filt_d205.mtx", soupX_objs[[1]]$out)
DropletUtils:::write10xCounts("gamm_soupX_filt_S2.mtx", soupX_objs[[2]]$out)
ref.matrix.soupx<- soupX_objs[[1]]$out
gamm.matrix.soupx<- soupX_objs[[2]]$out
```
# Create Seurat objects and convert to sce objects for doublet removal
```{r}
project_names <- purrr::map_chr(all_aligned_data, 'project')
feature_set1 <- list(feature1 = 'nCount_RNA', feature2 = 'nFeature_RNA')
filtered_list_of_objs <- purrr::map2(soupX_objs, project_names, ~create_seurat_and_sce(out = .x$out, project = .y, feature_set = feature_set1))

```
# Doublet removal and lane merging
```{r}
combine_feature_plots(filtered_list_of_objs, feature_set1, file_name_base = "post_soupx_qc")
filtered_sce_list <- list(filtered_list_of_objs[[1]]$sce, filtered_list_of_objs[[2]]$sce)
lane_and_merged_seurat_obj <- run_scDblFinder_and_merge(filtered_sce_list)
```
# get sparse count matrices
```{r}
dim(lane_and_merged_seurat_obj$seurat_obj1@assays$RNA@counts)
dim(lane_and_merged_seurat_obj$seurat_obj1@assays$RNA@data)
dim(lane_and_merged_seurat_obj$seurat_obj2@assays$RNA@counts)
dim(lane_and_merged_seurat_obj$seurat_obj2@assays$RNA@data)
all.equal(lane_and_merged_seurat_obj$seurat_obj2@assays$RNA@counts,
lane_and_merged_seurat_obj$seurat_obj2@assays$RNA@data)
all.equal(lane_and_merged_seurat_obj$seurat_obj1@assays$RNA@counts,
lane_and_merged_seurat_obj$seurat_obj1@assays$RNA@data)
ref.matrix.flt <- lane_and_merged_seurat_obj$seurat_obj1@assays$RNA@data
gam.matrix.flt <- lane_and_merged_seurat_obj$seurat_obj2@assays$RNA@data
# write out data
DropletUtils:::write10xCounts("reh_dblfinder_filt_d205.mtx", ref.matrix.flt)
DropletUtils:::write10xCounts("gamm_dblfinder_filt_S2.mtx", gam.matrix.flt)
rm(ref.matrix.soupx, gamm.matrix.soupx,lane_and_merged_seurat_obj)
```
# read in orthologs
```{r orthologs}
orthologs <- read.csv2("Human_Pig_Biomart_Filtered_mod.txt", sep="\t", header=TRUE)
```
# subset matrices
```{r subset}
# subset genes so they match only orthologous ones
human.orthos <- as.vector(orthologs$human.gene.name)
ref.matrix.flt<- ref.matrix.flt[rownames(ref.matrix.flt) %in% human.orthos, ]
# subset orthologs to they only match expr matrix
orthologs<- orthologs[orthologs$human.gene.name %in% rownames(ref.matrix.flt), ]
# re-subset raw matrix
human.orthos <- as.vector(orthologs$human.gene.name)
# check length to make sure they are the same
length(rownames(ref.matrix.flt))
length(orthologs$human.gene.name)
```
# switch to pig genes
```{r pig-genes}
# reorder based on exp matrix genes
ind_reorder <- match(rownames(ref.matrix.flt),orthologs$human.gene.name)
orthologs.reorder<- orthologs[ind_reorder,]
# now replace rownames of exp matrix with pig genes
pig.orthos <- as.vector(orthologs.reorder$pig.gene.name)
head(row.names(ref.matrix.flt))
row.names(ref.matrix.flt) <- pig.orthos
# check exp matrix row names
head(row.names(ref.matrix.flt))
length(rownames(ref.matrix.flt))
```
# load pig data and subset so both pig and human are the same
```{r pig-data}

# subset pig data so that there is same genes
dim(gam.matrix.flt)
gam.matrix.flt<- gam.matrix.flt[rownames(gam.matrix.flt) %in% pig.orthos, ]
dim(gam.matrix.flt)

# subset human refernce based on gamm data
ref.matrix.flt<- ref.matrix.flt[rownames(ref.matrix.flt) %in% rownames(gam.matrix.flt), ]
dim(ref.matrix.flt)
DropletUtils:::write10xCounts("ref.filtered.exp", ref.matrix.flt)
DropletUtils:::write10xCounts("gamm.data.filtered.exp", gam.matrix.flt)
#ref.matrix.flt <- Read10X("ref.filtered.exp/")
#gam.matrix.flt <- Read10X("gamm.data.filtered.exp/")
```
# filter out duplicated genes in reference
```{r filter}
# get rid of duplicates by averaging them
# Aggregate rows based on row names
ref.matrix.flt <- aggregate(ref.matrix.flt, by = list(rownames(ref.matrix.flt)), FUN = mean)
# use column 1 as rownames
rownames(ref.matrix.flt)<- ref.matrix.flt$Group.1
ref.matrix.flt <- select(ref.matrix.flt,  -c("Group.1"))
# turn back to sparse matrix
#install.packages("Matrix", repos="http://R-Forge.R-project.org")
library(Matrix)
ref.matrix.flt<- as.matrix(ref.matrix.flt)
ref.matrix.flt <- Matrix(ref.matrix.flt, sparse = TRUE)
```

# get metadata
```{r metadata}
# read in meta.data
metadata <- read.csv2("GSE142526_metadata/F6/cca_fetalvsorg_125CP_205_metadata.csv", sep=",", header=TRUE, row.names= 1)
# subset by D205 data
metadata <- subset(metadata, metadata$orig.ident=="D205")
metadata<- as.data.frame(metadata)
complete.cases(metadata)
metadata<-na.omit(metadata)

# get numbers for each celltype
table(metadata$type)
# drop AC1 because only 2
metadata <- subset(x = metadata, type != 'AC1')
# replace underscore with dash
row.names(metadata) <- sub("_1", "", row.names(metadata))
metadata.rows <-as.vector(row.names(metadata))
dim(metadata)
dim(ref.matrix.flt)
```
# create seurat object
```{r} 
human_D205.seurat<- CreateSeuratObject(ref.matrix.flt, project = "human_D205", assay = "RNA",
                                       min.cells = 3, min.features = 200)
human_D205.seurat<-AddMetaData(object=human_D205.seurat, metadata=metadata)
gamm.S2.seurat<- CreateSeuratObject(gam.matrix.flt, project = "gamm_S2", assay = "RNA",
                                       min.cells = 3, min.features = 200)
```
# save seurat object
```{r}
saveRDS(human_D205.seurat, file = "human_D205_ortholog.rds")
saveRDS(gamm.S2.seurat, file = "GAMM_S2_ortholog.rds")

```
# Mitochondrial gene filtering
```{r}
pig_mt_list <- c("ND1","ND2", "COX1", "COX2", "ATP8", "ATP6", "COX3", "ND3", "ND4L", "ND4", "ND5", "ND6", "CYTB")
human_D205.seurat_mt_filtered <- filter_cells(human_D205.seurat, mt.list = pig_mt_list)
gamm.S2.seurat_mt_filtered <- filter_cells(gamm.S2.seurat, mt.list = pig_mt_list)
```
# Normalization
```{r}
human_D205.seurat_norm <- normalize_data(human_D205.seurat_mt_filtered)
gamm.S2.seurat_norm <- normalize_data(gamm.S2.seurat_mt_filtered)
saveRDS(human_D205.seurat_norm, file = "human_D205_norm.rds")
saveRDS(gamm.S2.seurat_norm, file = "GAMM_S2_norm.rds")
```
## further steps: reference undergoes feature selection, scaling, and dimensional reduction, query does not
# Feature selection
```{r}
human_D205.seurat_feat_select <- feature_selection(human_D205.seurat_norm)
```

# Scaling data
```{r}
human_D205.seurat_scaled <- scale_data(human_D205.seurat_feat_select)
```

# Dimensional reduction
```{r}
human_D205.seurat_dim_reduced <- run_and_visualize_pca(human_D205.seurat_scaled)
```
# Run umap
```{r}
human_D205.seurat_umap <- run_umap(human_D205.seurat_dim_reduced)
saveRDS(human_D205.seurat_umap, file = "human_D205_umap.rds")
```