---
title: "seurat_mapping_for_cross_species"
author: "Beth Moore"
date: "2023-07-18"
output: html_document
---
# Seurat tansfer mapping for cross species analysis
# do preprocessing on human and pig data (preprocess_crossspecies.Rmd)
# both reference and query should undergo normalization and find variable features,
# then only reference is scaled and undegoes dimentionality reduction

# Load libraries, set wd, and source functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(patchwork)
library(scran)
library(BiocParallel)
library(DropletUtils)
library(cowplot)
library(harmony)
library(reticulate)
library(purrr)
library(jsonlite)
library(rmarkdown)
library(ggplot2)
library(scPred)
BETHS_GAMM_DATA_DIR <- "/w5home/bmoore/scRNAseq/GAMM/"
use_condaenv("scRNAseq_best")
setwd(paste0(BETHS_GAMM_DATA_DIR, "human_data/reh_cellrep_2020/"))
timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
output <- paste0("output_seurat_mapping_", timestamp)
dir.create(output, showWarnings = FALSE)
file.copy("/w5home/bmoore/scRNAseq_library/config.json", file.path(output, "config.json"))
config <- fromJSON(file.path(output, "config.json"))
output <- paste0(output, "/")
source("/w5home/bmoore/scRNAseq_library/src/sc_pipeline_functions.R")
```

# read in seurat objects that were preprocessed
```{r read_data}
gamms2<- readRDS(file = "output_seurat_20230718_115135/GAMM_S2_feat_select.rds")
human_D205.seurat<- readRDS(file = "output_20230711_155556/human_D205_umap.rds")
```
# prep for integration
```{r prep_for_integration}
ret.list<- list(gamms2,human_D205.seurat)
features <- SelectIntegrationFeatures(object.list = ret.list, nfeatures = 3000)
#ret.list <- PrepSCTIntegration(object.list = ret.list, anchor.features = features)
```
# reference umap plot
```{r ref_umap}
p2 <- DimPlot(human_D205.seurat, reduction = "umap", group.by = "type", label = TRUE, repel = TRUE) +
  NoLegend()
pdf(paste0(output, "human_ref-umap.pdf"), width = 8, height = 6)
print(p2)
dev.off()
```
## mapping and annotating query datasets
# get query data and find anchors
```{r get_query_data_add_anchors}
gamm.anchors <- FindTransferAnchors(reference = human_D205.seurat, query = gamms2,
                                        dims = 1:30, #reference.reduction = "pca",
                                        reduction= "rpca")
predictions <- TransferData(anchorset = gamm.anchors, refdata = human_D205.seurat$type,
                            dims = 1:30)
gamms2 <- AddMetaData(gamms2, metadata = predictions)
```
```{r save_query_data}
saveRDS(gamms2, file= paste0(output,"gamms2_rpca_pred.rds"))
```
# evaluate how well our predicted cell type annotations match the full reference. 
# need manual annotation data
```{r read_metadata}
# add in metadata from previous analysis
metadata.gamm <- read.csv("gamm_manual_annot_metadata_S2.txt", row.names = 1, 
            header = TRUE, sep = "\t")
# check metadata with query data
colnames(gamm@assays$RNA@data)[1:10]
rownames(metadata.gamm)[1:10]
# remove "_1" from metadata
rownames(metadata.gamm) <- sub("_1", "", rownames(metadata.gamm))
rownames(metadata.gamm)[1:10]
# add metadata to query data
gamms2 <- AddMetaData(gamms2, metadata.gamm)
```
# compare predictions to manual annotations
```{r compare_predictions_to_manual_annotations}
gamms2$prediction.match <- gamms2$predicted.id == gamms2$CellType
table(gamms2$prediction.match)
# examine some canonical cell type markers for specific cell populations.
table(gamms2$predicted.id)
table(gamms2$CellType)
```
# need marker genes to look at prediction score
#VlnPlot(pancreas.query, c("REG1A", "PPY", "SST", "GHRL", "VWF", "SOX10"), group.by = "predicted.id")
# projection of a query onto the reference UMAP structure.
```{r project_query_on_ref}
human_D205.seurat <- RunUMAP(human_D205.seurat, dims = 1:30, reduction = "pca", return.model = TRUE)
gamms2 <- MapQuery(anchorset = gamm.anchors, reference = human_D205.seurat, query = gamms2,
                           refdata = list(celltype = "type"), reference.reduction = "pca", reduction.model = "umap")
```
# visualize
```{r visualize}
p1 <- DimPlot(human_D205.seurat, reduction = "umap", group.by = "type", label = TRUE, label.size = 3,
              repel = TRUE) + ggtitle("Reference annotations")# + NoLegend()
p2 <- DimPlot(gamms2, reduction = "ref.umap", group.by = "predicted.celltype", label = TRUE,
              label.size = 3, repel = TRUE) + ggtitle("Query transferred labels")
pdf(paste0(output, "celltype_seurat_predictions-umap2.pdf"), width = 11, height = 6)
print(p1 + p2)
dev.off()
p3 <- DimPlot(gamms2, reduction = "ref.umap", group.by = "CellType", label = TRUE,
              label.size = 3, repel = TRUE) + ggtitle("Manually annotated labels")
pdf(paste0(output, "celltype_seurat_predictions-umap_c0.14.pdf"), width = 11, height = 6)
print(p2 + p3)
dev.off()
```
# visualize probabilities
```{r visualize_prob}
#  probabilities of each cell in the @meta.data slot of the query Seurat object.
# get labels
labels<- as.vector(unique(colnames(gamms2@meta.data)))
labels<-labels[! labels %in% c('orig.ident', 'nCount_RNA', 'nFeature_RNA', 
        'predicted.id','sizeFactors','scDblFinder.score','percent.mt','CellType',
        'ident','scDblFinder.weighted', 'prediction.score.max','prediction.match', 
        'RNA_snn_res.0.14','RNA_snn_res.0.5',"predicted.celltype.score","predicted.celltype",
        'scDblFinder.class','scDblFinder.cxds_score','seurat_clusters')]
# visualize the probabilities over the UMAP plot:
pdf(paste0(output, "celltype_prediction_featureplot2.pdf"), width = 10, height = 10)
print(FeaturePlot(gamms2, labels, keep.scale = "all"))
dev.off()
```
# get comparison between predicted and manual annotations
```{r get_comparison}
# verify model performance in query data
query.ref.data.table<- as.data.frame(crossTab(gamms2, "CellType", "predicted.celltype"))
# check out proportion of cells
query.ref.data.table2<- as.data.frame(crossTab(gamms2, "CellType", "predicted.celltype", output = "prop"))
write.table(query.ref.data.table, file = paste0(output, "query_ref_data_table_s2.txt"), sep = "\t")
write.table(query.ref.data.table2, file = paste0(output, "query_ref_data_table_proportion_s2.txt"), sep = "\t")
```